\documentclass[11pt,letterpaper]{article}
\usepackage{amssymb,amsfonts,color,graphicx,amsmath,enumerate}
\usepackage{tikz}
\usepackage{amsthm}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\Lp}[2]{\left\|{#1}\right\|_{L^{#2}}}
\newcommand{\field}{\mathbb{F}}
\newcommand{\affine}{\mathbb{A}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\text{Var}}
\newcommand{\ind}{\mathbbm{1}}
\newcommand{\Cov}{\text{Cov}}

\newenvironment{solution}
{\begin{proof}[Solution]}
{\end{proof}}

\voffset=-3cm
\hoffset=-2.25cm
\textheight=24cm
\textwidth=17.25cm
\addtolength{\jot}{8pt}
\linespread{1.3}

\begin{document}
\begin{center}
{\bf \Large Math 130B - Expectation}
\vspace{0.2cm}
\hrule
\end{center}

\begin{enumerate}
	\item Let $X$ and $Y$ be jointly continuous random variables.
	\begin{enumerate}
		\item Show that if $X$ and $Y$ are independent, then $\E[XY] = \E[X]\E[Y]$.
		\vfill

		\item Give an example of two jointly continuous random variables $X$ and $Y$ for which $\E[XY] \neq \E[X]\E[Y]$.
	\end{enumerate}

	\vfill

	\item Consider the continuous iid random variables $T_1, \ldots, T_n$. For any $k = 1, 2, \ldots, n$, find
	\[
		\E\left(\frac{T_1 +T_2 + \cdots + T_k}{T_1 + T_2 + \cdots + T_n}\right).
	\]

	\vfill

	\item Let $G = (V,E)$ be a graph with $n$ vertices and $e$ edges. Show that you can partition the vertices of $G$ into two subsets $A$ and $B$ in such a way that half of the edges in $G$ cross between $A$ and $B$.

	\vfill

	\item Let $\sigma$ be a random permutation of $\{1, 2, \ldots, n\}$, chosen uniformly. Let $X$ be the number of fixed points of $\sigma$ (recall that $i$ is a fixed point of $\sigma$ if $\sigma(i) = i$). Find $\E[X]$.

	\vfill

	\item A \emph{tournament} with $n$ players is a special kind of directed graph. Imagine all $n$ players play two-person games with each other. Draw an arrow from player $u$ to player $v$ if player $u$ beat player $v$ in their game (assume there are no draws).

	A \emph{path} is a sequence of edges $v_1 \to v_2 \to \ldots v_k$ where $v_i \to v_{i+1}$ is a directed edge and the vertices are all distinct. The path is called \emph{Hamiltonian} if it contains all vertices in the graph.

	Show that there is a tournament $T$ with $n$ players and at least $n!\cdot 2^{-(n-1)}$ Hamiltonian paths.

	\vfill
\end{enumerate}

\end{document}