\documentclass[12pt]{article}  

\pagestyle{empty}

\usepackage{graphics}
\usepackage{amsmath,amssymb,amsthm, multicol,enumerate}
\usepackage[pdftex]{graphicx}
\usepackage{epsf}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\Aut}{\text{Aut}}
\newcommand{\Lp}[2]{\left\|{#1}\right\|_{L^{#2}}}
\newcommand{\tr}{\text{tr}}
\newcommand{\field}{\mathbb{F}}

\addtolength{\oddsidemargin}{-.75in}
\addtolength{\evensidemargin}{-.75in}
\addtolength{\textwidth}{1.5in}
\addtolength{\topmargin}{-1in}
\addtolength{\textheight}{2.25in}

\begin{document}
\begin{center}
{\bf \Large Math 13 - Week 1: Formal Logic}
\vspace{0.2cm}
\hrule
\end{center}

\begin{enumerate}
	\item What do you think the point of this class is? What do you think we're going to cover? How do you think this class fits into your math education? Do you think you'll enjoy this class?

	\vfill

	\item Write the following statement in if-then form: ``Every integer divisible by 5 ends in a 0 or a 5 (when written in its usual decimal form).''

	\vfill

	\item Write the converse of the statement ``If you love me, then you will marry me.''

	\vfill

	\item Which of the following statements are true and which are false? You don't have to prove your answers (but you should give it a shot!)
	\begin{enumerate}
		\item Every integer is positive or negative.
		\item Every integer is even and odd.
		\item If $x$ is an integer and $x>2$ and $x$ is prime, then $x$ is odd.
		\item Let $x$ and $y$ be integers. We have $x^2 = y^2$ if and only if $x=y$.
		\item The sides of a triangle are all congruent to each other if and only if its three angles are all $60^\circ$.
		\item If an integer $x$ satisfies $x = x+1$, then $x=6$.
	\end{enumerate}

	\vfill

	\item Are the expressions $x\implies \neg y$ and $\neg (x\implies y)$ logically equivalent?

	\vfill

	\item Is the expression $(x\implies y)\lor (x\implies \neg y)$ a tautology (true for all values of $x$ and $y$)?
	\vfill

	\item Can you write $x\implies y$ using only $\neg$, $\lor$ and $\land$?
	\vfill
\end{enumerate}


\end{document}