\documentclass[11pt]{article}
\usepackage{amsmath, amssymb, amsthm, verbatim,enumerate,bbm, mathtools}
\usepackage{indentfirst}



\date{\today}
\parindent 5mm
\parskip 0.2mm
\oddsidemargin  0pt \evensidemargin 0pt \marginparwidth 0pt
\marginparsep 0pt \topmargin 0pt \headsep 0pt \textheight 8.8in
\textwidth 6.6in
\renewcommand{\baselinestretch}{1.03}

\allowdisplaybreaks

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{property}[theorem]{Property}
\newcommand{\Bin}{\ensuremath{\textrm{Bin}}}
\newcommand{\Unif}{\ensuremath{\textrm{Unif}}}
\newcommand{\Var}{\ensuremath{\textrm{Var}}}
\newcommand{\Cov}{\ensuremath{\textrm{Cov}}}
\newcommand{\Exp}{\ensuremath{\textrm{Exp}}}
 
 
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\F}{\mathbb{F}}



\title{Midterm -- Math 130B}
\author{Instructor: Liam Hardiman}
\date{\today}

\begin{document}

\maketitle


{\bf Instructions: } \begin{itemize} 
\item You must show your work and clearly
explain your line of reasoning.
%\item You may use a calculator, but only for basic arithmetic. If you didn't bring a calculator, you may use your phone, but it must be set to airplane/flight mode and you must clear all notifications before you start.
\item You have one hour and twenty minutes to complete the exam.
\end{itemize}

\medskip
\medskip

\center{{\bf GOOD LUCK!}}

\newpage
\ 
\newpage

\begin{enumerate}
    \item (10 points)
    Suppose the measurement errors, $X_i$, from $n$ experiments are uniformly distributed on $[-1, 1]$.
    Assume that the errors are independent.
    \begin{enumerate}
        \item (4 points)
        Find the probability distribution function for $|X_i|$.

        \vfill

        \item (6 points)
        Find the probability density function and probability distribution functions for the largest absolute error.
        Be sure to clearly state the domains of the PDF and CDF.
        \emph{[Hint: consider $Y = \max(|X_1|, \ldots, |X_n|)$.]}

        \vfill
    \end{enumerate}

    \newpage
    \ 
    \newpage

    \item (10 points)
    Let $X$ and $Y$ be independent random variables with $X\sim \Exp(\lambda)$ and $Y\sim \Unif [a,b]$.
    Let $U = 2X-3Y$ and $V = 4X+Y$.
    Find the joint probability density function for $(U, V)$.
    Be sure to state the domain of this function.

    \newpage
    \ 
    \newpage

    \item (10 points)
    Let $X\sim \mathcal{N}(0, 1)$ and $Y\sim \mathcal{N}(0, 1)$ be independent.
    Let $Z = X/Y$.
    \begin{enumerate}
    \item (5 points)
    Find the conditional probability density $f_{Z\mid Y}(z\mid Y=y)$.
    \vfill

    \item (5 points)
    Find the joint probability density $f_{Z, Y}(z, y)$ and use this to find the marginal density $f_Z(z)$.
    What type of random variable is $Z$?
    \vfill

    \end{enumerate}

    \newpage
    \ 
    \newpage

    \item (10 points)
    Let $X$ and $Y$ be random variables with joint density
    \[
        f(x,y) = \begin{cases}
            ce^{-(x+y)},&\text{if }0<x<y<\infty\\
            0,&\text{otherwise.}
        \end{cases}
    \]
    \begin{enumerate}
        \item (2 points)
        Find the value of $c$ such that $f$ is a density function.

        \vfill

        \item (5 points)
        What is the conditional density of $X$ given that $Y = y$.
        Be sure to state the domain of this function.

        \vfill

        \item (3 points)
        Are $X$ and $Y$ independent? Why or why not?
        \vfill
    \end{enumerate}

    \newpage
    \ 
    \newpage

    \item (10 points)
    True or false?
    Give a brief justification for each answer.

    \begin{enumerate}
        \item Suppose $X$ is a continuous random variable with distribution function $F_X$. 
        Then there exists a random variable $Y$ with distribution $F_Y$ satisfying $F_Y(y) = 2F_X(y)$.

        \vfill

        \item Let $X$ and $Y$ be random variables such that $E[XY] = E[X]E[Y]$.
        Then $X$ and $Y$ are independent.

        \vfill

        \item If $X$ and $Y$ are independent random variables, then $E[XY] = E[X]E[Y]$.

        \vfill

        \item If $X$ and $Y$ are independent random variables, then $\Var[X+Y] = \Var[X] + \Var[Y]$.
        \vfill

        \item If the events $E_1$ and $E_2$ are independent and the events $E_2$ and $E_3$ are independent, then the events $E_1$ and $E_3$ are independent.

        \vfill
    \end{enumerate}

    \newpage
    \ 
    \newpage
    \ 

\end{enumerate}


\end{document}
