\documentclass[11pt,letterpaper]{article}
\usepackage{amssymb,amsfonts,color,graphicx,amsmath,enumerate}
\usepackage{tikz}
\usepackage{amsthm}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\Lp}[2]{\left\|{#1}\right\|_{L^{#2}}}
\newcommand{\field}{\mathbb{F}}
\newcommand{\affine}{\mathbb{A}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\text{Var}}
\newcommand{\ind}{\mathbbm{1}}
\newcommand{\Cov}{\text{Cov}}

\newenvironment{solution}
{\begin{proof}[Solution]}
{\end{proof}}

\voffset=-3cm
\hoffset=-2.25cm
\textheight=24cm
\textwidth=17.25cm
\addtolength{\jot}{8pt}
\linespread{1.3}

\begin{document}
\begin{center}
{\bf \Large Math 130B - Midterm Review Problems}
\vspace{0.2cm}
\hrule
\end{center}

\begin{enumerate}
	\item A model proposed for NBA basketball supposed that when two teams with roughly the same record play each other, the number of points scored in a quarter by the home team minus the number scored by the visiting team is approximately a normal random variable with mean 1.5 and variance 6.
	Additionally, the model supposes that the point differentials for the four quarters are independent.
	Suppose we are working within this model.
	\begin{enumerate}
		\item What is the probability that the home team wins?
		\vfill

		\item What is the conditional probability that the home team wins, given that it is behind by 5 points at halftime.
		\vfill

		\item What is the conditional probability that the home team wins, given that it is ahead by 5 points at the end of the first quarter?
	\end{enumerate}
	\vfill

	\item Let $X_{(1)} \leq X_{(2)} \leq \cdots \leq X_{(n)}$ be the ordered values of $n$ independent uniform $(0,1)$ random variables.
	\begin{enumerate}
		\item Prove that for $1\leq k \leq n+1$,
		\[
			\Pr[X_{(k)} - X_{(k-1)} > t] = (1-t)^n,
		\]
		where $X_{(0)} \equiv 0$ and $X_{(n+1)} \equiv 1$.
		\vfill

		\item What is the distribution for the median of the sample? \textit{That is, find the distribution for the median of $X_{(n)} - X_{(1)}$}.
	\end{enumerate}

	\vfill\null\pagebreak

	\item Suppose $X$ is a random variable taking only nonnegative integer values. Show that if $\E[X] < 1$, then $\Pr[X = 0] > 0$.

	\vfill

	\item Suppose $X$ and $Y$ are jointly continuous random variables with joint pdf given by
	\[
		f(x,y) = ce^{-(x+y)},\quad 0\leq y\leq x <\infty
	\]
	\begin{enumerate}
		\item Find the value of the constant $c$.
		\vfill

		\item Find the conditional pdf of $Y$ conditioned on $X=x$.
		\vfill

		\item Are $X$ and $Y$ independent?
		\vfill

		\item Suppose $U = X+2Y$ and $V = X-Y$. Find the joint pdf of $U$ and $V$. Be sure to specify the domain!
		\vfill
	\end{enumerate}
\end{enumerate}

\end{document}