\documentclass[11pt,letterpaper]{article}
\usepackage{amssymb,amsfonts,color,graphicx,amsmath,enumerate}
\usepackage{tikz}
\usepackage{amsthm}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\Lp}[2]{\left\|{#1}\right\|_{L^{#2}}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\affine}{\mathbb{A}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\text{Var}}
\newcommand{\ind}{\mathbbm{1}}
\newcommand{\Cov}{\text{Cov}}

\newenvironment{solution}
{\begin{proof}[Solution]}
{\end{proof}}

\voffset=-3cm
\hoffset=-2.25cm
\textheight=24cm
\textwidth=17.25cm
\addtolength{\jot}{8pt}
\linespread{1.3}

\begin{document}
\begin{center}
{\bf \Large Math 173A - Midterm Practice}
\vspace{0.2cm}
\hrule
\end{center}

\begin{enumerate}

    \item \begin{enumerate}
        \item What is the order of 2 in $\F_{31}$?
        \item Is 3 a primitive root in $\F_{31}$?
    \end{enumerate}


    \item Let $g$ be a primitive root of $\F_p$, where $p$ is a prime.
    Show that $g^k$ is a primitive root of $\F_p$ if and only if $\gcd(k, p-1) = 1$.

    \item Alice and Bob agree to use the prime $p=71$ and the base $g=7$ for a Diffie-Hellman key exchange.
    \begin{enumerate}
        \item Alice chooses $a = 12$ as her secret and Bob chooses $b = 31$ as his secret.
        What are the values $A$ and $B$ that they should send to each other?

        \item What is their shared secret?
    \end{enumerate}

    \item Use the baby step giant step algorithm to find $\log_2(15)$ in $\F_{29}$.

    \item \begin{enumerate}

    \item Describe each step in the ElGamal cipher.

    \item If Eve has an oracle that solves the discrete logarithm problem, can she decrypt Elgamal ciphertexts? How?
    \end{enumerate}

    \item \begin{enumerate}
        \item Define what $f(x) = O(g(x))$ means.
        \item Is $e^{x^2} = O(e^x)$?
        \item Is $e^x = O(e^{x^2})$?
        \item Is $\ln(x) = O(\log x)$?
    \end{enumerate}

    \item Let $p$ and $q$ be primes and suppose that $p = 2q+1$.
    Let $h\in \F_p^\times$ and assume that $h\neq 1$.
    Prove that $x^2 \equiv h\pmod p$ has a solution if and only if the order of $h$ is $q$.

\end{enumerate}

\end{document}