\documentclass[11pt,letterpaper]{article}
\usepackage{amssymb,amsfonts,color,graphicx,amsmath,enumerate}
\usepackage{tikz}
\usepackage{amsthm}

\newcommand{\naturals}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\complex}{\mathbb{C}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\mcal}[1]{\mathcal{#1}}
\newcommand{\rationals}{\mathbb{Q}}
\newcommand{\Lp}[2]{\left\|{#1}\right\|_{L^{#2}}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\affine}{\mathbb{A}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\text{Var}}
\newcommand{\ind}{\mathbbm{1}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\Pois}{\text{Pois}}
\newcommand{\Bin}{\text{Bin}}
\newcommand{\Bern}{\text{Bern}}


\newenvironment{solution}
{\begin{proof}[Solution]}
{\end{proof}}

\voffset=-3cm
\hoffset=-2.25cm
\textheight=24cm
\textwidth=17.25cm
\addtolength{\jot}{8pt}
\linespread{1.3}

\begin{document}
\begin{center}
{\bf \Large Math 130B - Homework 4}
\vspace{0.2cm}
\hrule
\end{center}



\begin{enumerate}

    \item Let $X_1, X_2, \ldots$ be a sequence of independent random variables with identical distribution. Let their mean be $\mu$ and their variance be $\sigma^2$.
    Define $S_n = X_1 + \cdots + X_n$.
    For any $a > 0$, compute
    \[
        \lim_{n\to \infty}\Pr[\mu \leq S_n \leq \mu + a].
    \]

    \item Let $X$ be a geometric random variable with parameter $p$.
    \begin{enumerate}
        \item Compute the moment generating function, $M(t)$ of $X$.
        Be sure to specify for which values of $t$ the moment generating function is finite.

        \item Use the moment generating function to compute the mean and variance of $X$.
    \end{enumerate}

    \item You have a lottery ticket. There are four possible outcomes, each of which has probability 1/4 of occurring.
    \begin{enumerate}[(i)]
        \item You win nothing;
        \item You win \$5 and another lottery ticket;
        \item You win two lottery tickets;
        \item You win \$10.
    \end{enumerate}
    Compute the expected value of your total winnings. Assume that the lottery tickets that can be won in outcomes (ii) and (iii) have the same properties as the one you started with.

    \item Let $X$ and $Y$ be independent uniform random variables distribute don $[0,3]$ and $[1,3]$. What is the probability that $Y > X^2-1$?

    \item Consider two components and three types of electrical shock.
    A type 1 shock causes component 1 to fail.
    A type 2 shock causes component 2 to fail, and a type 3 shock causes both components 1 and 2 to fail.
    The times until shocks 1, 2, and 3 occur are independent exponential random variables with respective rates $\lambda_i$ for $i = 1,2,3$.
    Let $X_i$ denote the time at which component $i$ fails.
    Find $\Pr[X_1 > s, X_2 > t]$.

    \item Let $f(x,y) = \frac{e^{-x/y}e^{-y}}{y}$ for $x,y\geq 0$.
    Compute the probability that $X>1$ given that $Y = y$ for some $y$.

    \item \begin{enumerate}
        \item Let $X$ be a discrete random variable whose possible values are $1, 2, \ldots$.
        If $\Pr[X = k]$ is nonincreasing in $k$, prove that
        \[
            \Pr[X = k] \leq 2\frac{E[X]}{k^2}.
        \]

        \item Let $X$ be a nonnegative continuous random variable having a nonincreasing density function.
        Show that
        \[
            f(x) \leq \frac{2E[X]}{x^2},\quad\text{for all }x>0.
        \]
    \end{enumerate}

    \item Prove or provide a counterexample.
    \begin{enumerate}
        \item If $X$ and $Y$ are two independent random variables with the same distribution, then $\Cov(X+Y, X-Y) = 0$.

        \item If $X_1, X_2, \ldots$ is a sequence of independent and identically distributed random variables with $E[X_i] = 0$, then
        \[
            \lim_{n\to \infty}\Pr\left[\frac{1}{n}|X_1 + \cdots + X_n| \leq 0.01\right] = 1.
        \]
    \end{enumerate}


    \item Let $x_i\geq 0$ for $1\leq i \leq n$ and assume $x_1 + \cdots + x_n = 1$.
    Prove that
    \[
        -\sum_{i=1}^nx_i \ln(x_i) \leq \ln n.
    \]
    \textit{Hint: Introduce a random variable $X$ which takes the values $x_1, \ldots, x_n$ each with probability $1/n$.}

    \item Let $X$ and $Y$ be independent random variables with $X\sim \Pois(\lambda)$ and $Y\sim \Bern(p)$ with $0<p<1$.
    Define the random variable $Z = XY$.
    \begin{enumerate}
        \item Determine the probability mass function $p_Z$ of $Z$.
        \item Find the moment generating function $M_Z(t)$ of $Z$.
        \item Compute the covariance $\Cov(Y, Z)$ and correlation $\rho(Y, Z)$.
    \end{enumerate}

\end{enumerate}

\end{document}